\label{chap:acknowledgements}

\begin{center}
  \uppercase{Acknowledgements}
  \addcontentsline{toc}{section}{Acknowledgements} % insert into TOC
\end{center}

\bigskip

% *****************************************************************

I grew up on a farm, where my parents boarded and briefly bred horses. 
We put up hay, I learned to garden at eight.  We were paid in pop to fill the loft. 
I didn't know what McDonald's was until second grade.
We had a big driveway where my brothers and I played street hockey, and where I learned to fail.
My mother was a waitress and drove a paper route for over two decades each.
My father was a forklift operator when I was very young, and retired from the US Postal Service in Spring Lake, MI this year after 26 years of service.
 
I lost my brother Cody on May 19, 2007 when I was 20 and he 19;
his girlfriend Emily Rankin gave birth to my nephew Landon on November 30 of that year.
Cody went to university before I did.
I bet he would have been the first of our whole family to get a college degree before me, too.
I wish he could have seen Grampa's eyes that day.
My older brother Eric died on January 31, 2009. 
He was 25.
Eric also had an Emily.  
She married a hockey player years after he passed and gave birth to triplet boys this year.
I cherish every moment I can spend with my brother, Montana, and he has grown to be a great man.

I worked my way through undergrad at the One Trick Pony at first a dishwasher, and then saut\'e.  
Wiley dropped a sushi knife on his foot.  Santos had a big mustache and a bigger comb.  I love those guys.
I met Aaron Schutza through our friend Will, and I will always admire his mind.  He was the first person to hug me when Eric died.
We both wrote our senior theses under Milun Racovi\'c on separate years, and I'll always be grateful for your intuition and rigor and kindness, Milun.

Karen Gipson got me into WSU to pursue a PhD in physics.  I met and lived and worked with Brett Deaton there.  
He was the first person to hug me when my parents announced they were to divorce.  
%I love that man like a brother.
We studied under Matthew Duez, though I came into the group three years after he did.
I have never known someone with such a command of the English language and astrophysics.  This manuscript bears his and Francois Foucart's imprint.
Abigail Higgins is my best friend.
There is no more loving, friendly and charming companion than she.




% *****************************************************************

\newpage