\chapter{Equation of State Simulations of Black Hole-Neutron Star Mergers}
\label{chap:chapter-4}

In this chapter, we review the literature of numerical relativity in the context of black hole-neutron star mergers, primarily focusing on finite-temperature, nuclear-theory based equations of state.  
A robust general relativity code is needed not only to model the gravitational waveform, but also to produce the electromagnetic signals that can be detected by infrared or optical instruments.  
In this sense, the coalescence of black hole neutron star binaries and their mergers are essential systems that can be probed with multi-messenger astronomy.   
There are two essential components of the fluid matter: (1) the tidal tail formed during the merger of the two objects and (2) the remnant accretion disk left after the merger.  
Both of which can produce optically bright outflows.

(1) A large amount of matter in the tail inevitably becomes unbound from the gravitational potential of the black hole.  Matter is said to be ``unbound'' if $u_t < -1$, where $u_alpha$ are the covariant components of the fluid 4-velocity.  
The disruption causes unbound matter to be ejected from the system, expanding, cooling, and forming heavy elements via r-process nucleosynthesis.
These unstable nuclei then undergo radioactive decay and fission on the timescale of days after the merger, i.e. ``kilonova'', which can be observed in the near infrared spectrum and on the timescale of a week.
Meanwhile, the gravitational wave signal may only be detectable on the timescale of inspiral and merger ($\sym10$ ms).
The chemical composition of the tail material $Y_e$ provides a measurement of the amount of heavy elements created during the nucleosynthesis of the ejecta. 
Extremely neutron-rich matter ($Y_e \le 0.25$) is expected to produced 2nd and 3rd, but not 1st, r-process elements.  
Their cosmogenic origin of the r-process elements is still uncertain today, so getting a sense of how many of the heavy elements are produced during the merger can help us address this problem.
  
(2)  The accretion disk also contributes to the outflow via disk winds (e.g. by viscous or neutrino-absorption heating), although the disk ejecta typically is a factor of a few smaller compared to the tail.
However, disk outflows occur only about 100 ms later than the nucleosynthesis in the tail, and can produce electromagnetic transients via interaction with the interstellar gas, emitting late radio afterglow.
One difference between the two that might be significant is that disk outflows are expected to undergo more neutrino processing and be more neutron starved (higher $Y_e$), which would affect what elements are formed.
Likewise, the polar regions of the disk outflow (jets) will also contribute to the total optical signal output of the event in the form of GRBs.

Hot equations of state were first simulated in Newtonian physics~\cite{ruff1996-leakage_part1,ruff1997-leakage_part2,ruff2001-leakage_part3,ross2002-leakage_part1,ross2003-leakage_part2,ross2003-leakage_part3}.
They found accretion dynamics were strongly dependent on the equation of state, where stiffer equations of state did not undergo a mass transfer in a single pass, but occurred more episodically.
By modifying the Newtonian method with a pseudo-potential that emulates general relativity by artificially creating an ISCO, they determined that there was no discrepancy with compaction on disruption. 
Later,~\cite{Duez:2009yy} used the general relativity-based code and found that disruption always happens in a single event, even with a stiff equation of state.
Thus, Newtonian simulations are unreliable.
The following sections are summarizing work in the field of black hole-neutron star mergers using general relativity, so we will not discuss Newtonian simulations any further.

% Kyoto group fitiing formulae paper
\section{BHNS binaries with zero-temperature, piecewise polytrope EOS}

Before reviewing the simulations of hot nuclear equations of state, 
we first discuss the findings of the Kyoto group where mixed binaries are studied with a large swath of piecewise polytrope zero-temperature equations of state with the adaptive mesh refinement numerical relativity code, \SACRA (SimulAtor for Compact objects in Relativistic Astrophysics).  
As mentioned in Chapter \ref{chap:chapter-2}, the utility of a PP equation of state is that it provides an approximation to realistic equations of state, making it easier to systematically explore a small number of different parameter estimates for the core and crust material of the star.  

In~\cite{Kyutoku:2013}, four different equations of state with a wide range of compactnesses $\mathcal{C} = (0.138 - 0.180)$ were studied with black hole spin $\chi = (0, 0.5, 0.75)$, and mass ratio $q \equiv M_{\rm BH}/M_{\rm NS} = (3, 5, 7)$ with the initial configuration for the star always assuming an ADM mass of $M_{\rm NS} = 1.35 M_\odot$.
This survey led to the estimation of a semi-analytic fitting formula to determine the mass and velocity of the dynamical ejecta~\cite{Kyutoku:2015}.  They obtained:
\begin{align}
\label{e:vej_sacra}
\left\langle v_{\rm ej} \right\rangle = (0.01533 q + 0.1907) c
\end{align}
An analytic model for the light curve of the resulting kilonova/macronova is also determined.

The resulting mass of the ejecta is determined to be:
\begin{align}
\label{e:mass_ejecta}
\frac{M_{\rm ej}}{M^b_{\rm NS}} = 
{\rm Max} 
\left \{ 
a_1 q^{n_1} (1- 2 \mathcal{C})\mathcal{C}^{-1}
-a_2 q^{n_2} \bar{r}_{\rm ISCO}(\chi_{\rm eff} )
+a_3 \left( 1- \frac{M_{\rm NS}}{M^b_{\rm NS}} \right)
+a_4
,
0
\right \},
\end{align}
where the neutron star has compactness $\mathcal{C}$, baryon and ADM mass $M^b_{\rm NS}$ and $M_{\rm NS}$, respectively, and the mixed binary system has mass ratio $q$.
The effective spin of the black hole is determined by the tilt, $i$ between the spin $\chi$ and the orbital angular momentum and ($\chi_{\rm eff} = \chi \cos i$)
The radius of the ISCO is determined by solving the formula
\begin{align}
\label{e:isco_r}
\bar{r}_{\rm ISCO}(\chi) &=
3+Z_2
\mp (\chi)
\sqrt{(3 - Z_1) (3 + Z_1 + 2 Z_2)} \\
\label{e:isco_z1}
Z_1 &= 
1 + (1 - \chi^2)^{1/3}
\left \{ 
(1 + \chi)^{1/3} + (1 - \chi)^{1/3}
\right \}  \\
\label{e:isco_z2}
Z_2 &= \sqrt{3 \chi^2 + Z_1^2}.
\end{align}
The fitting paremeters $\{n_i,a_i\}$ were determined from their simulation data via the method of least squares.

Prior to this study, ~\cite{Foucart2012} determined a semi-analytic formula to relate the dimensionless parameters of the initial configuration system ($\mathcal{C}_{\rm NS}, q, \chi_{\rm BH} $) with the baryonic mass of the post-merger remnant:
\begin{align}
\label{e:mass_disk}
\frac{M^b_{\rm rem}}{M^b_{\rm NS}} = 
\alpha (3 q)^{1/3} (1 - 2 \mathcal{C} )
-\beta  \frac{R_{\rm ISCO}}{R_{\rm NS}},
\end{align}
where the fitting parameters here are $\alpha = 0.288$ and $\beta = 0.148$.
This formula was fitted to reported disk masses available numerical relativity literature.
One should note that $\bar{r}_{\rm ISCO} = R_{\rm ISCO}/M_{\rm BH}$ in the ejecta mass prediction is the radius of the ISCO in (dimensionless)  Boyer-Lindquist form.
The determined remnant mass $M^b_{\rm rem}$ is defined to be any bound material outside the black hole 10 ms after merger.  
Much of that matter isn't quite disk or protodisk yet, but fallback.

Lastly, Pannarele determined semi-analytical fitting formulae for the final (10 ms after merger) black hole mass, spin, and inclination angle~\cite{PannaraleEtAl2013,Pannarale:2014}.
The black hole mass formula is based on the Buonanno, Kidder, and Lehner (BKL) approach of double black hole mergers to estimate the final spin, which itself accounts for first-order mass conservation, and that energy and angular momentum radiate away during inspiral~\cite{Buonanno2008a}.  
In the presence of a neutron star, the baseline BKL approach is augmented by taking into account the (predicted) mass of the remnant disk (via Foucart's prediction formula \ref{e:mass_disk}), and further extended to include gravitational wave emission.
Once the final spin magnitude  $\chi_f$ and inclination angle $i_f$ of the black hole is determined from a two-dimensional root-solving scheme, the final black hole mass is calculated to yield
\begin{align}
M^f_{\rm BH} &= 
M \left \{ 
1 - \left[1-e(\bar{r}^i_{\rm ISCO}, \chi_i) \right]
\right \}
- e(\bar{r}^f_{\rm ISCO}, \chi_f) M^b_{\rm rem}\\
e(r, \chi) &= 
\frac{ r^2 - 2 r \pm \chi \sqrt{r} }
{ r \left(r^2 - 3r \pm 2 \chi \sqrt{r}\right)^{1/2} }
\end{align}
where $M$ is the total system mass, $M = M_{\rm BH} +M_{\rm NS}$

% Foucart et al (2014)
\section{BHNS Binaries with LS220 Equation of State}

While the inspiral of these systems can be adequately modeled with a simpler, approximate equation of state like a piecewise polytrope, the disruption phase requires a more detailed equation of state with composition and temperature dependence to more fully describe the physical effects.  
To produce the late time gravitational wave signal, when finite size effects of the neutron star are potentially detectable and point mass post-Newtonian approximations begin to have significant errors, one must also evolve the system in the last tens of orbits prior to disruption.  
On the other hand, if the effort is to predict the electromagnetic signal and the nucleosynthesis yields, the number of requisite orbits can be relaxed.  
In addition to the dependencies of neutron star structure (temperature, composition, pressure), we also have to accurately model magnetic field effects of the post merger remnant disk and neutrino radiation on the disrupting fluid (and as a significant source of cooling on the disk).
Nuclear reactions weren't modeled, since they were in nuclear statistical equilibrium, but during merger non-$\beta$-equilibrium parts of the table were sampled.

Neutrinos can also play a role on the composition of the unbound material ($u_t < -1$) that gets ejected from the system, as well as the heavy r-process elements created there during disruption.
This is quite significant in double neutron star mergers, but probably not for mixed binaries~\cite{Roberts:2016}.
Neutrino-antineutrino annihilation in the less dense regions and near the poles of the black hole can contribute to the production of short gamma ray bursts.  
While magnetic fields and the magneto-rotational instability (MRI) are crucial to the late-stage evolution of the accretion disk, which contribute to disk winds via magnetically driven outflows ~\cite{Paschalidis2014,Kiuchi:2015qua}.
The equation of state, however, plays a significant role in all stages of evolution: compared to a stiff equation of state, the relative response of a soft star to the tidal disruption from the black hole during inspiral occurs closer to the black hole (likewise, if at all), can lead to qualitatively different features during merger (thinner stream of matter falling into the black hole), and the multivariate dependence of the fluid evolution on the equation of state are all essential.

In~\cite{Foucart:2014nda}, the first relativistic simulations of black hole-neutron star mergers with a finite temperature equations of state with an accompanying neutrino leakage scheme were performed.  
This work used the Lattimer-Swesty equation of state discussed in Section \ref{sec:ls220}, and explored the parameter space where the mass of the neutron star is in the lower mass range ($M_{\rm NS} = 1.2 M_\odot - 1.4 M_\odot$), while the black hole mass is in the peak of the distribution of likely stellar-mass black holes ($M_{\rm BH} = 7 M_\odot - 10 M_\odot$). 
For these lower-mass black holes, a larger spin is necessary to produce maximal disruption ($\chi_{\rm BH} = 0.7 - 0.9$)~\cite{Foucart2012}.  
Their survey was also restricted in the sense that magnetic fields were not included since the disk wasn't evolved for long times (thermal timescales of the disk), noting that the effects of neutrino cooling on the inner regions of the disk and the viscous heating due to the MRI turbulence might more or less offset each other.

Comparing to the piecwise polytropes of~\cite{hebeler2013equation}, it was found that the merger dynamics using their hot model were less discernible compared to cold models for systems involving lower mass black holes.  
For higher mass black holes, on the other hand, the difference was most notable as the more rapid disruption occurred later and closer to the black hole with a much thinner tail than with the commonly used (in simulations) $\Gamma = 2$ polytrope (with the same configuration parameters).

There were several limitations noted at the time the simulations were performed.
First, the grid stricture of the finite difference grid was not adaptable at the time.  
That is, given that the tail is much larger in structure than the disk, accurately resolving the accretion disk would lead to over-resolving the tail plus evolving enormous amounts of vacuum--far too computationally expensive.
Conversely, lowering resolution by allowing the grid to expand in order to track the ejecta in the tail would drastically under resolve the disk.  
Therefore, to evolve the disk, the large amount of unbound matter had to exit the grid unfortunately early.  Second, a full study of the disk evolution would require the effects of MRI turbulence, or an equivalent effective viscosity routine.
The neutrino leakage scheme used in their work--as well as ours--only goes as far as an estimate for the presence of neutrinos (no transport, absorption effects or spectral information).
A more robust neutrino transport scheme would be needed to follow the production of neutrino driven winds and electron-positron creation in the baryon-poor polar axes of the black hole (gamma ray bursts).

Key findings from this study found that a finite temperature equation of state neutron stars reliably ejects a large amount of material ($M_\textrm{ej} = 0.04 M_\odot -  0.16 M_\odot$), where less massive stars eject more material, and form accretion disks with mass $M_\textrm{disk} = 0.05 M_\odot -  0.15 M_\odot$, where larger black hole spin lend to more massive disks (and therefore less bound tail material).
The disks are initially hot ($T_\textrm{disk} \sim 5-15 \textrm{MeV}$) and bright in neutrinos ($L_\nu \sim 10^{53} \textrm{erg/s}$).
In the first 10 ms after merger, the disk quickly protonizes from $Y_e \sim 0.06$ to $Y_e \sim 0.1 - 0.4$ because the electron antineutrino emission is highest.  
At later times, $Y_e$ falls because of disk cooling.
They also found recoil velocities from the asymmetric ejection of matter that were much larger than the kick velocities imparted by gravitational waves.

% Deaton et al (2013)
\subsection{BHNS post-merger remnant study of LS220 Equation of State}

In a more in-depth study of one of the models studied in that paper, ~\cite{Deaton2013} examined the effects of neutrino cooling on the remnant disk with the configuration system $M_{\rm NS} = 1.4 M_\odot$, $M_{\rm BH} = 5.6 M_\odot$, $\chi_{\rm BH} = 0.9$ for the LS220 equation of state.  
This was the first study of a finite-temperature, composition-dependent equation of state black hole-neutron star system with full general relativity and neutrino cooling.  
Although their disk remnant was only evolved to a fraction of the $\sym 200$ms accretion timescale, on the thermal timescale of $\sym 40$ms, as beyond that the long-term effects of not using magnetic fields would render the simulation unphysical.
Magnetic fields should drive accretion via angular momentum transport from magnetic winding and the magnetorotational instability (MRI)-induced turbulence.
In their simulation, accretion only happens because the disk is settling and circularizing, heating only via shocks.
Once it starts to axisymmetrize, there is no more angular momentum transport so the fluid stops accreting. 
However, there are two sources of cooling: neutrino emission and the advection of hot matter onto the black hole, with the former dominating at late times in nonmagnetized cases.

In this study, the post merger remnant had baryon rest mass $M_\textrm{disk} \sim 0.3 M_\odot$, much larger than the disk masses in the~\cite{Foucart:2014nda} study.  
This is to be expected because of the smaller black hole mass.
The neutrino luminosity peaked $\sym 10$ ms after merger at $L_\nu \sim 10^{54} \textrm{erg} s^{-1}$ and fell to a fifth of this value 50 ms after merger---the disk was quite optically thick in the neutrino sense.  
The temperature of the disk and the resulting structure was susceptible to a substantial amount of neutrino cooling, which makes the areal radius of the disk smaller than hotter remnants but still hot and optically thick.
Neutrino cooling first drives up $Y_e$ as emission is antineutrino-dominant (causing more neutrons to turn into protons)
and then the equilibrium $Y_e$ shifts back down as the gas cools.  
Electrons become degenerate since there are fewer positrons for $n + e^+ \rightarrow p + \bar{\nu_e}$.

The ejecta had a mass of $M_{\rm ej} \sim 0.08 M_\odot$ with an average velocity $\sim 0.2 c$, although some matter was ejected at $0.5 c$. 
Additionally, the tail, while under resolved in their study, was very neutron rich with (density averaged) chemical composition $\left\langle Y_e \right\rangle_\textrm{ej} \sim 0.1$  and in nuclear statistically equilibrium as the temperature remained $T_\textrm{ej} \sim 1 \textrm{MeV}$.  
This was the only black hole-neutron star study that found a second source of ejecta, resulting from the shock region near the black hole, instead of only the tail.
This secondary ejecta source is more common in double neutron star binaries, since this second component is spilled away where the stars collide, and provides a more protonization, greater entropy component of the ejecta, but not enough to significantly affect nucleosynthesis outcomes, as found in ~\cite{Roberts2012}.
The observable radio transient in the interaction between the ejecta and interstellar medium was estimated to be within the detectable range of the Expanded Very Large Array (EVLA).


% Foucart et al (2016)
\section{Precessing BHNS binaries with hot, microphysical EOS}

Another recent exploration in the parameter space of hot, microphysical black hole-neutron star systems was done in~\cite{FoucartDD2:2017}.
This was the first time such mergers were studied in the parameter space in which the black hole spin was inclined with respect to the orbital angular momentum. 
In these simulations we used the DD2 equation of state (cf. Section \ref{sec:dd2}), again varying black hole mass $M_{\rm BH} = (5, 7) M_\odot$ and spin $\chi_{\rm BH} = (0.7, 0.9)$ with neutron star mass of $M_{\rm NS} = 1.4 M_\odot$, but this time not restricting the highly prograded spin of the black hole to be orbitally aligned.
The inclination angle was set to $i_{\rm BH}=60^{\circ}$ in all but the high-spin high-mass case where we set $i_{\rm BH}=20^{\circ}$, since the ejecta is only expected for high-mass black holes for nearly-aligned mergers.
We note here that the tidal tail is again much narrower than for the ideal gas or piecewise polytrope equation of state.  

The ejecta was measured to fall in the range of $M_{\rm ej} \sim (0.01 - 0.05) M_\odot$, while higher mass black holes and higher spinning black holes causing the most ejecta.  
However, the ejecta masses were much smaller than that of aligned-spin black holes in the previous study~\cite{Foucart:2014nda}: the magnitude of the orbitally aligned component of the black hole spin is the central parameter in driving ejecta and tidal disruption.  
The unbound material was again very neutron rich with $\left\langle Y_e \right\rangle_\textrm{ej} \sim 0.04 - 0.06$, so that heavy elements and high-opacity lanthanides will be produced.
All of the material has composition $Y_e \le 0.25$ such that near this upper limit lighter r-process elements will be produced.

As a note, the fitting parameters for the unbound mass used in this work (Version 2 on \url{arxiv.org}) were different to the parameters that are now available to the public (Version 3 on \url{arxiv.org}).
The calculation of the ISCO radius (Equations \ref{e:isco_r} - \ref{e:isco_z2}) was also different, using instead the innermost stable \textit{spherical} orbit (ISSO) from Pannarelle~\cite{Pannarale:2014}, which is used in both the disk- and ejecta-mass fitting formulae (Equations \ref{e:mass_disk} and \ref{e:mass_ejecta}, respectively).

Comparing to the fitting formulae in Equation \ref{e:vej_sacra}, the results of the simulations with the hot, nuclear-theory based equation of state were quite different.  
For the $5 M_{\rm BH}$ black hole system, the fitting formulae predict that that 
$\left\langle v_{\rm ej, rms} \right\rangle = 0.245 c$.
The result in the \SpEC simulations gave $\left\langle v_{\rm ej, rms} \right\rangle = 0.175 c$.
To account for the $~40 \%$ difference between the two results, one has to keep in mind that the fitting formula of  Equation \ref{e:vej_sacra} was based on the use of zero-temperature equations of state, no composition dependence and bet-equilibrium.
The finite-temperature DD2 equation of state, on the other hand, only assumes muclear statistic equilibrium and is composition dependent.  

There is a physical difference: the effective chemical composition for the cold, $\beta$-equilibrium equation of state at low-densities is $Y_e \sim 0.5$ (mostly $^{56}{\rm Ni}$), while the neutrino leakage calculation ~\cite{Ruffert1996,Rosswog:2003rv,OConnor2010} on the DD2 equation of state simulations produced $Y_e \sim 0.05$ (mostly neutrons and $\sim 10\%$ seed nuclei at cold temperatures) since the composition mostly advects.
If all of the energy required to transform from neutron-rich ejecta to $^{56}{\rm Ni}$ was turned into kinetic energy, the simulations would produce ejecta with average velocity $\left\langle v_{\rm ej, rms} \right\rangle = 0.22 c$, within $\sim 10\%$ of the fitting formula prediction.
Advection of $Y_e$ is more realistic at early times, but at later times will beta-decay toward higher $Y_e$, so some of this energy input actually happens, although not when it appears in the \SACRA simulations.
If half of the energy released by beta-decays produces relativistic electrons and neutrinos, $\sym 4.6 {\rm MeV/nuc}$ of kinetic energy can be accounted for in comparison to the piecewise polytrope models.
We get the updated fitting formula:
\begin{align}
\label{e:vej_spec}
\left\langle v_{\rm ej, rms} \right\rangle = (0.0166 q + 0.1667) c
\end{align}


After merger (where half of the baryonic matter has already been accreted onto the black hole), the temperature of the fluid reaches a clear local minimum, coinciding with a minimizing accretion rate.  Shortly after, the tail begins to self-intersect, causing hydrodynamic shocks that reheat the fluid and causes the bound material to begin circularizing around the black hole.  
This protodisk begins to axisymmetrize $\sim 10$ ms after merger.  
For the M5-S7-I60 system, the disk reached a finite average temperature of $3 - 6$ MeV and average electron fraction $Y_e \sim 0.1 - 0.2$.  The disk, still accreting at a rate of $\dot{M}_{\rm BH} \sim 5 M_\odot {\rm s}^{-1}$, has a mass of $M_{\rm disk} = (0.045 - 0.06) M_\odot$ 17 and 10 ms after merger.

In some of the \SpEC simulations in the past, the ejecta was ran through a smooth particle hydrodynamics (SPH) code coupled to the SkyNet nuclear reaction network to determine nucleosynthesis output~\cite{Lippuner2015,Roberts:2016}.
It was found in that work 2nd and 3rd peaks in the r-process, as expected, with a minor production of 1st peak elements seeded by late-time ($\sym100$ms) irradiation of ejecta by neutrinos from the remnant. 